---
title: "Which composite indicator for which data?"
author: "Romain Lorrilliere"
date: "2023-06-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE,error = FALSE)
```
```{r source}
library(data.table)
library(ggplot2)
library(glmmTMB)
library(ggeffects)
library(scales)

source("c:/git/LPI_theoritical_test/functions/fun_dyn_pop.r")
source("c:/git/LPI_theoritical_test/functions/fun_samples.r")
source("c:/git/LPI_theoritical_test/functions/fun_lpi.r")
source("c:/git/LPI_theoritical_test/functions/fun_composite_indicators.r")

```
# Purpose
There are many kinds of composite indicators to aggregate dynamic population survey data. 
The data available to calculate these composite indicators can vary widely. They can vary over gradients of quality and quantity, for example, by the number of populations monitored, by the number of years monitored, by the completeness of the monitoring (number of years monitored over the number of years between the first and last year) and by the sampling effort (or method) affecting the probability of detection. 


This theoretical work defines which composite indicators are the most efficient according to the data. 


# Methods

## Virtual population dynamics

To get data representing simple to complex scenarios, we model the following cases: 


We start with very simple cases

* 1 species 1 pop
* 1 species 2 pops
* 2 species of 1 pop
* 2 species from 2 pops

Furthermore, to get data closer to natural conditions, we will also need to run scenarios with more species and more populations: 50 species with some populations between 10 and 50 


We use several dynamics scenario 

* Stability
* all species and pop increase or decline
  + Uniform
  + Proportional (extinction after n time steps)
  + Exponential (with carrying capacity)

More complex cases, scenarios with compensation: 
Compensation can be observed in abundance, where the sum of gains and losses equals zero, or in variations in abundance, where the product of variations is equal to 1. Moreover, compensation can be balanced between species, 1/2 increasing and 1/2 declining (50/50), or unbalanced 90/10 and 75/25. 

Finally we have two more parameters to test : 

* Without stochasticity, temporal and between populations
* With initial abundances, large or small (25, 50, 100)


```{r param_sim}

ny <- 50
nb_pop_sync <- 10
nb_rep_random <- 30

```



### One species with one population with uniform dynamics

```{r dyn}
tparam <- expand.grid(model_name = "uniform",N0 = c(10,50,100), K = Inf,r_mean = c(5,0,-5),demographic_stocha = "none",nb_pop = 1)
setDT(tparam)
tparam[,`:=`(id_sc = paste0("1_",1:.N),nb_sp = 1)]

d1 <- f_make_dyn_species(model_name = "uniform",demographic_stocha = "none",r_mean = tparam$r_mean,N0_mean = tparam$N0,K = tparam$K,nb_pop = nrow(tparam),nb_year = ny)


setnames(d1,c("r","K"),c("r_pop","K_pop"))
d1[,`:=`(id_sc = pop, iteration = 1, id_iteration = paste0(pop,"_i1"))]

setkey(tparam,id_sc)
setkey(d1,id_sc)
d1 <- d1[tparam]
d1[,`:=`(sp = 1,pop = 1, id_sp = paste0(id_iteration,"_s",1),id_pop = paste0(id_iteration,"_s",1,"_p",1),sc_name = "uniforme",K_sp = K_pop,r_sp = r_pop,
         sync_cos_div = NA, sync_cos_time =  NA, sync_cos_power = NA , sync_sin_div = NA, sync_sin_time = NA, sync_sin_power = NA)]

theCol <- c("id_sc","sc_name","model_name","demographic_stocha","sync_cos_div", "sync_cos_time" , "sync_cos_power"  , "sync_sin_div" , "sync_sin_time" , "sync_sin_power", "nb_sp","nb_pop" ,"N0","K","r_mean","iteration","id_iteration","id_sp", "sp", "K_sp","r_sp", "id_pop" ,"pop" ,"K_pop","r_pop" ,"year","N")

d1 <- d1[,theCol,with=FALSE]

d <- copy(d1)


```

### One species with one population with exponential dynamics

```{r dyn2}
tparam <- expand.grid(model_name = "logistic",N0 = c(10,50,100), K = Inf,r_mean = c(0.1,0,-0.1),demographic_stocha = "none",nb_pop = 1)
setDT(tparam)
tparam[,K := N0 * 10]
tparam[,`:=`(id_sc = paste0("1_",1:.N),nb_sp = 1)]


d1 <- f_make_dyn_species(model_name = "logistic",demographic_stocha = "none",r_mean = tparam$r_mean,N0_mean = tparam$N0,K = tparam$K,nb_pop = nrow(tparam),nb_year = ny)


setnames(d1,c("r","K"),c("r_pop","K_pop"))
d1[,`:=`(id_sc = pop)]


setkey(tparam,id_sc)
setkey(d1,id_sc)
d1 <- d1[tparam]
d1[,id_sc := paste0(2,substr(id_sc,2,nchar(id_sc)))]
d1[,`:=`(iteration = 1, id_iteration = paste0(id_sc,"_i1"))]


d1[,`:=`(sp = 1,pop = 1, id_sp = paste0(id_iteration,"_s",1),id_pop = paste0(id_iteration,"_s",1,"_p",1),sc_name = "logistic",K_sp = K_pop,r_sp = r_pop,
         sync_cos_div = NA, sync_cos_time =  NA, sync_cos_power = NA , sync_sin_div = NA, sync_sin_time = NA, sync_sin_power = NA)]


theCol <- c("id_sc","sc_name","model_name","demographic_stocha","sync_cos_div", "sync_cos_time" , "sync_cos_power"  , "sync_sin_div" , "sync_sin_time" , "sync_sin_power", "nb_sp","nb_pop" ,"N0","K","r_mean","iteration","id_iteration","id_sp", "sp", "K_sp","r_sp", "id_pop" ,"pop" ,"K_pop","r_pop" ,"year","N")

d1 <- d1[,theCol,with=FALSE]


d <- rbind(d,d1)




```
One species with several populations with exponential dynamics and temporal variations

We create a less linear temporal series by adding a periodic process to get a more realistic data set. Here the temporal variations have two different sources. The growth rate could vary synchronously between populations or species (as observed in CES data). So, first, we add a complex periodic function to compute a part of the annual growth rate. Then, the temporal stochasticity corresponds to picking the abundance in a Poisson distribution at each time.


First the periodic process



```{r sync_param}
tparam <- expand.grid(model_name = "logistic",N0 = c(10,50,100), K = Inf,r_mean = c(0.1,0,-0.1),demographic_stocha = "none",nb_pop = nb_pop_sync)
setDT(tparam)
tparam[,K := N0 * 10]
tparam[,`:=`(id_sc = paste0("1_",1:.N),nb_sp = 1)]

sc <- 1
it <- 1


idsc <- paste0("3_",sc)
idit <- paste0(idsc,"_i",it)
idsp <- paste0(idit,"_s1")
```


```{r sync, echo=TRUE}

d_sync <- data.table(year = seq(from=1,to= ny, by = 1))

d_sync[,`:=`(id_iteration = idit, p_cos_div = round( runif(1,0,1),2), p_cos_time = round( runif(1,1,4),1), p_cos_power = round(runif(1,1,3),2),
             p_sin_div = round( runif(1,0,1),2), p_sin_time = round( runif(1,1,4),1), p_sin_power = round(runif(1,1,3),2))]


d_sync[, r_periodic := round(rescale(cos(year *  p_cos_div) + 
                                       cos(year *p_cos_time) + 
                                       cos(year ^ p_cos_power) + 
                                       sin(year *  p_sin_div) + 
                                       sin(year *p_sin_time) + 
                                       sin(year^p_sin_power),
                                     to=c(-0.25,0.25)),
                             3)]
print(head(d_sync))
r_sync<- d_sync[,r_periodic]

d_sync_all <- copy(d_sync)

```

This process creates an additional value of growth rate balanced around 0. 

The summary : 


```{r sync_fig}

print(summary(r_sync))

gg <- ggplot(data= d_sync,aes(x=r_periodic)) + geom_histogram() + labs(x = "periodic additionnal effect on growht rate")
print(gg)




gg <- ggplot(data=d_sync, aes(x = year, y = r_periodic))+ geom_line() + geom_point()
eq_txt <- paste0("cos(",d_sync$p_cos_div[1]," * t) + cos(",d_sync$p_cos_time[1]," * t) + cos(t^",d_sync$p_cos_power[1],") + sin(",d_sync$p_sin_div[1]," * t) + sin(",d_sync$p_sin_time[1]," * t) + sin(t^",d_sync$p_sin_power[1],")")
gg <- gg +  labs(x = "year", y = "periodic additionnal effect on growth rate",title = idit, subtitle = eq_txt)

print(gg)



```





Then, at each time step, these values of the periodic effect of growth rate are now added to the growth rate, and we simulate `r nb_pop_sync` population abundance trajectories.


```{r dyn_sync}


di <-  f_make_pop_dyn(nb_pop = nb_pop_sync, N0_mean = tparam$N0[sc],N0_var = 0,K=tparam$K[sc],K_method="exact",r_mean = tparam$r_mean[sc], r_init_var = 0,r_temporal_var_mean = 0,r_synchro = r_sync,nb_year = ny,model_name = "logistic",demographic_stocha = "poisson", fig="print",format_output = "table")

setDT(di)
setnames(di,c("r","K"),c("r_pop","K_pop"))


di[,`:=`(id_sc =idsc,sc_name = "logistic_sync",model_name = "logistic",demographic_stocha = "poisson",
             sync_cos_div = d_sync$p_cos_div[1], sync_cos_time =  d_sync$p_cos_time[1], sync_cos_power = d_sync$p_cos_power[1] ,
             sync_sin_div = d_sync$p_sin_div[1], sync_sin_time = d_sync$p_sin_time[1], sync_sin_power = d_sync$p_sin_power[1],
             nb_sp = 1, nb_pop = tparam$nb_pop[sc], N0 = tparam$N0[sc],K = tparam$N0[sc],r_mean = tparam$r_mean[sc],
             iteration = it,id_iteration = idit, id_sp = idsp, sp = 1, id_pop = paste0(idsp,"_p",pop),K_sp = K_pop,r_sp = r_pop)]

theCol <- c("id_sc","sc_name","model_name","demographic_stocha","sync_cos_div", "sync_cos_time" , "sync_cos_power"  , "sync_sin_div" , "sync_sin_time" , "sync_sin_power", "nb_sp","nb_pop" ,"N0","K","r_mean","iteration","id_iteration","id_sp", "sp", "K_sp","r_sp", "id_pop" ,"pop" ,"K_pop","r_pop" ,"year","N")

di <- di[,theCol,with=FALSE]


d <- rbind(d,di)

for(sc in 2:nrow(tparam)) {
  
  di <-  f_make_pop_dyn(nb_pop = nb_pop_sync, N0_mean = tparam$N0[sc],N0_var = 0,K=tparam$K[sc],K_method="exact",r_mean = tparam$r_mean[sc], r_init_var = 0,r_temporal_var_mean = 0,r_synchro = r_sync,nb_year = ny,model_name = "logistic",demographic_stocha = "poisson", fig="",format_output = "table")

setDT(di)
setnames(di,c("r","K"),c("r_pop","K_pop"))

idsc <- paste0("3_",sc)
idit <- paste0(idsc,"_i",it)
idsp <- paste0(idit,"_s1")
di[,`:=`(id_sc =idsc,sc_name = "logistic_sync",model_name = "logistic",demographic_stocha = "poisson",
             sync_cos_div = d_sync$p_cos_div[1], sync_cos_time =  d_sync$p_cos_time[1], sync_cos_power = d_sync$p_cos_power[1] ,
             sync_sin_div = d_sync$p_sin_div[1], sync_sin_time = d_sync$p_sin_time[1], sync_sin_power = d_sync$p_sin_power[1],
             nb_sp = 1, nb_pop = tparam$nb_pop[sc], N0 = tparam$N0[sc],K = tparam$K[sc],r_mean = tparam$r_mean[sc],
             iteration = it,id_iteration = idit, id_sp = idsp, sp = 1, id_pop = paste0(idsp,"_p",pop),K_sp = K_pop,r_sp = r_pop)]

theCol <- c("id_sc","sc_name","model_name","demographic_stocha","sync_cos_div", "sync_cos_time" , "sync_cos_power"  , "sync_sin_div" , "sync_sin_time" , "sync_sin_power", "nb_sp","nb_pop" ,"N0","K","r_mean","iteration","id_iteration","id_sp", "sp", "K_sp","r_sp", "id_pop" ,"pop" ,"K_pop","r_pop" ,"year","N")

di <- di[,theCol,with=FALSE]


d <- rbind(d,di)
  
  
}


```


```{r sync_loop}

for(it in 2:nb_rep_random) {
  
  idit <- paste0(idsc,"_i",it)
idsp <- paste0(idit,"_s1")

d_sync <- data.table(year = seq(from=1,to= ny, by = 1))

d_sync[,`:=`(id_iteration = idit,p_cos_div = round( runif(1,0,1),2), p_cos_time = round( runif(1,1,4),1), p_cos_power = round(runif(1,1,3),2),
             p_sin_div = round( runif(1,0,1),2), p_sin_time = round( runif(1,1,4),1), p_sin_power = round(runif(1,1,3),2))]


d_sync[, r_periodic := round(rescale(cos(year *  p_cos_div) + 
                                       cos(year *p_cos_time) + 
                                       cos(year ^ p_cos_power) + 
                                       sin(year*  p_sin_div) + 
                                       sin(year *p_sin_time) + 
                                       sin(year ^p_sin_power),
                                     to=c(-0.25,0.25)),
                             3)]

r_sync<- d_sync[,r_periodic]

d_sync_all <- rbind(d_sync_all,d_sync)


for(sc in 1:nrow(tparam)) {
  
  di <-  f_make_pop_dyn(nb_pop = nb_pop_sync, N0_mean = tparam$N0[sc],N0_var = 0,K=tparam$K[sc],K_method="exact",r_mean = tparam$r_mean[sc], r_init_var = 0,r_temporal_var_mean = 0,r_synchro = r_sync,nb_year = ny,model_name = "logistic",demographic_stocha = "poisson", fig="",format_output = "table")

setDT(di)
setnames(di,c("r","K"),c("r_pop","K_pop"))

idsc <- paste0("3_",sc)
idit <- paste0(idsc,"_i",it)
idsp <- paste0(idit,"_s1")
di[,`:=`(id_sc =idsc,sc_name = "logistic_sync",model_name = "logistic",demographic_stocha = "poisson",
             sync_cos_div = d_sync$p_cos_div[1], sync_cos_time =  d_sync$p_cos_time[1], sync_cos_power = d_sync$p_cos_power[1] ,
             sync_sin_div = d_sync$p_sin_div[1], sync_sin_time = d_sync$p_sin_time[1], sync_sin_power = d_sync$p_sin_power[1],
             nb_sp = 1, nb_pop = tparam$nb_pop[sc], N0 = tparam$N0[sc],K = tparam$K[sc],r_mean = tparam$r_mean[sc],
             iteration = it,id_iteration = idit, id_sp = idsp, sp = 1, id_pop = paste0(idsp,"_p",pop),K_sp = K_pop,r_sp = r_pop)]

theCol <- c("id_sc","sc_name","model_name","demographic_stocha","sync_cos_div", "sync_cos_time" , "sync_cos_power"  , "sync_sin_div" , "sync_sin_time" , "sync_sin_power", "nb_sp","nb_pop" ,"N0","K","r_mean","iteration","id_iteration","id_sp", "sp", "K_sp","r_sp", "id_pop" ,"pop" ,"K_pop","r_pop" ,"year","N")

di <- di[,theCol,with=FALSE]


d <- rbind(d,di)
  
  
}

  
}


```

```{r plot_all_sync}

gg <- ggplot(d_sync_all,aes(x=year,y=r_periodic))+ facet_wrap(.~id_iteration)
gg <- gg + geom_line()
gg

```

```{r sum_trajectories}
nbsc <- length(unique(d$id_sc))
nbit <- length(unique(d$id_iteration))


```

Finally we get `r nbit` trajectories of `r nbsc` different scenarios. 


```{r trajectories}


gg <- ggplot(data = d, aes(x =year, y = N, group = id_pop)) + geom_line(alpha = 0.6) + facet_wrap(.~id_sc,scale="free_y")
gg

```

## species abundance variations


For each species, we could aggregate the population abundances in different ways : 

 - the median and various quantile
 - the arithmetic mean 
 - the geometric mean
 
 
```{r model_glm_trend}
d_glm_trend_init <- data.frame(id_iteration = unique(d[,id_iteration]),trend=99.99,sd=99.99,ICinf=99.99,ICsup=99.99)

vec_serie <- c("all","start_10","mid_10","end_10")

d_glm_trend <- d_glm_trend_init
colanmes(d_glm_trend_s) <- paste(colnames(d_glm_trend_init,"all",sep="_"))
                                   

for(s in vec_serie[2:4]) {
  d_glm_trend_s <- d_glm_trend_init
  colanmes(d_glm_trend_s) <- paste(colnames(d_glm_trend_init,s,sep="_"))
  d_glm_trend <- merge(d_glm_trend,d_glm_trend_s,by="id_iteration")
} 


setDT(d_glm_trend)
for(i in d_glm_trend[,id_iteration]) {
  cat("\n",i,": ")
  for(s in vec_serie) {
    cat(s," ")
    if(s == "all") {
  di <- d[id_iteration == i,]
    } else if(s == "start_10") {
      di <- d[id_iteration == i & year < 11,]
    } else if(s =="mid_10") {
      di <- d[id_iteration == i & year >= (mean(year)-5 & year <= mean(year) + 5),]
    } else {
      di <- d[id_iteration == i & year >= max(year) - 10,]
    }
    
  nbpop <- di[1,nb_pop]
  if(nbpop == 1 ){
    #md for only one pop
    md <- try(glm(N~year, data = di, family= "quasipoisson"),silent = TRUE)  
    if(class(md)[1] != "try-error") {
      if(nbpop == 1 ){
      smd <- as.data.frame(summary(md)$coefficients) 
      } else {
        smd <- as.data.frame(coef(summary(md))$cond)
      }
      
      trend <- exp(smd[2,1])
      sd <- exp(smd[2,2])
      ci <- try(confint(md),silent = TRUE) 
      if(class(md)[1] != "try-error") {
         ICinf <- exp(ci[1,2])
      INsup <- exp(ci[2,2])
      } else {
         ICinf <- NA
      INsup <- NA
      }
    } else {
       trend <- NA
      sd <- NA
      ICinf <- NA
      INsup <- NA
    }
  d_glm_trend[id_iteration == i, paste0("trend_",s) := trend]
  d_glm_trend[id_iteration == i, paste0("sd_",s) := sd]
  d_glm_trend[id_iteration == i, paste0("ICinf_",s) := ICinf]
  d_glm_trend[id_iteration == i,paste0("ICsup_",s) := ICsup]

      
    }
  } 

}
  
 fwrite(d_glm_trend,"output/trend_ref.csv")
 print(head(d_glm_trend))
 
 
```
 
 
```{r model_gam_tot}
for(i in unique(d[,id_iteration])) {
  di <- d[id_iteration == i,]
  nbpop <- di[1,nb_pop]
  if(nb_pop ==1 ){
    md <- gam(N~year, data = di, family= poisson)  
    
  }
  
  
  
}


```


```{r summmary} 
theCol_pop <- c("id_sc","sc_name","model_name","demographic_stocha","sync_cos_div", "sync_cos_time" , "sync_cos_power"  , "sync_sin_div" , "sync_sin_time" , "sync_sin_power", "nb_sp","nb_pop" ,"N0","K","r_mean","id_iteration","id_sp", "id_pop")

dparam_pop <- unique(d[,theCol_pop,with=FALSE])

theCol_sc <- c("id_sc","sc_name","model_name","demographic_stocha","nb_sp","nb_pop" ,"N0","K","r_mean")

dparam_sc <- unique(d[,theCol_sc,with=FALSE])

dudu <- d[,theCol_sc,with=FALSE]
d_Nsp <- d[,.(N_tot=sum(N),N_med = median(N),N_25 = quantile(N,.25), N_75 = quantile(N,.75), N_Ar_mean = round(mean(N),2), N_Geom_mean = round(exp(mean(log(N+N0*0.01))),2)),by = .(id_sc, id_iteration,year, sc_name)]

print(head(d_Nsp))


dl_Nsp <- melt(d_Nsp,id.vars = c("id_sc","id_iteration","year","sc_name"),mesure.vars = c("N_tot","N_med","N_25","N_75","N_Ar_mean","N_Geom_mean"))
gg <- ggplot(data = dl_Nsp, aes(x =year, y = value, group = id_iteration,colour = sc_name)) + geom_line(alpha = 0.8) + facet_wrap(.~variable,scale = "free_y")
gg

```

Then we ave several way to assess the indices of abundance variations : 
Firstly with the previously aggregated data
 - Bray-Curtis
 - log-linear model with year as factor (glm)
 - additif model (GAM)
 
 

```{r N0}

d_Nsp_0 <- d_Nsp[year == 1 ,.(id_iteration,N_tot,N_med,N_25,N_75,N_Ar_mean,N_Geom_mean)]

var_names <- c("N_tot","N_med","N_25","N_75","N_Ar_mean","N_Geom_mean")

setnames(d_Nsp_0,var_names, paste0(var_names,"_0"))
setkey(d_Nsp_0,id_iteration)

setkey(d_Nsp,id_iteration)

d_Nsp <- merge(d_Nsp,d_Nsp_0)





```

```{r dN}


for(v in var_names){
  N <- d_Nsp[,v,with = FALSE]
  N0 <- d_Nsp[,paste0(v,"_0"),with=FALSE]
  d_Nsp[,X := N/N0]
  setnames(d_Nsp,"X",paste0("dN_",v))
}
  


```



```{r dN_analysis}
d_dN <- d_Nsp[,.(id_sc,id_iteration,year,dN_N_tot,dN_N_med,dN_N_25,dN_N_75,dN_N_Ar_mean,dN_N_Geom_mean)]
d_dN <- melt(d_dN,id.vars = c("id_sc","id_iteration","year"))
d_dN <- merge(d_dN,dparam_sc,by="id_sc")
setDT(d_dN)
d_dN_log <- d_dN[model_name != "uniform",]
d_dN_log[,id_combi := paste(N0,K,sep="_")]

setorder(d_dN_log,id_iteration,year)
gg <- ggplot(d_dN_log,aes(x = year, y = value, group = id_iteration, colour = variable)) + geom_line() + facet_grid(id_combi~r_mean,scales = "free")
print(gg)

```







```{r abundance_variation_bray-curtis}


for(v in var_names){
  N <- d_Nsp[,v,with = FALSE]
  N0 <- d_Nsp[,paste0(v,"_0"),with=FALSE]
  d_Nsp[,X := get_braycurtis_tot(N,N0)]
  setnames(d_Nsp,"X",paste0("BC_",v))
}
  



```




