---
title: "Generate the temporal variation of population abundance"
author: "Romain Lorrilliere"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE,error = FALSE)
```

```{r source}
source("c:/git/LPI_theoritical_test/functions/fun_dyn_pop.r")
```

# One population abundance

## Exponential dynamic

We generate the abundance variation of one population following an exponential function. 
Thus the population abundance at t time N_t is only according to :
N(t+1) = exp(r) * N(t)
N0

```{r the_funciton}
f_recursive_dyn_pop
```
```{r parm1}
r <- 0.05
N <- 100
demographic_stocha <- ""

```

For example :
r = `r r` and N0 = `r N`

```{r dyn1}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r, demo_var= demographic_stocha, fig="print")


```


```{r parm2}
r <- -0.05

```

or another example :
r = `r r` and N0 = `r N`

```{r dyn2}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r,demo_var= demographic_stocha, fig="print")


```
## Temporal variability

Then we add temporal variability in the r parameter.



```{r parm3}
r <- 0.05
r_var <- 0.025

```

For example :
r in N(mean=`r r`,sd = `r r_var`) and N0 = `r N`

```{r dyn3}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r, r_temporal_var = r_var,demo_var= demographic_stocha,fig="print")


```


```{r parm4}
r <- -0.05


```

or another example :
r in N(mean=`r r`,sd = `r r_var`) and N0 = `r N`

```{r dyn4}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r, r_temporal_var = r_var,demo_var= demographic_stocha,fig="print")


```

## Temporal stochasticity
One could also add a demographic stochasticity process by picking abundance from a Poisson distribution. 




```{r parm5}
r <- -0.05
demographic_stocha <- "poisson"

```

with the same paramters:
r in N(mean=`r r`,sd = `r r_var`) and N0 = `r N`

```{r dyn5}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r, r_temporal_var = r_var,demo_var= demographic_stocha,fig="print")


```

# Several populations

Finally, we can generate the dynamics of as many populations as necessary


```{r parm6}
r <- -0.05
N_var <- 10
r_var0 <- 0.02
nb_pop <- 5
```

for example for `r nb_pop` populations:

N0 in N(mean=`r N`,sd = `r N_var`)
and r in N(mean=N(mean = `r r`,sd = `r r_var0`),sd = `r r_var`) 

```{r dyn6}
l <- f_make_pop_dyn(nb_pop = nb_pop, N0_mean = N,N0_var = N_var,r_mean = r, r_init_var = r_var0, r_temporal_var = r_var,demo_var= demographic_stocha,fig="print")


```
