---
title: "Generate the temporal variation of population abundance"
author: "Romain Lorrilliere"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE,error = FALSE)
```

```{r source}
source("c:/git/LPI_theoritical_test/functions/fun_dyn_pop.r")
```

# One population abundance

## Exponential dynamic

We generate the abundance variation of the population following an exponential dynamics. 
Thus the population abundance at t time N_t is only according to :
N(t+1) = exp(r) * N(t)
N0

```{r the_funciton}
f_recursive_dyn_pop
```
```{r parm1}
r <- 0.05
N <- 100
```

For example :
r = `r r` and N0 = `r N`

```{r dyn1}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r, fig="print")


```


```{r parm2}
r <- -0.05
N <- 100
```

or another example :
r = `r r` and N0 = `r N`

```{r dyn2}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r, fig="print")


```

Then we add temporal variability in the r parameter.



```{r parm3}
r <- 0.05
r_var <- 0.1
N <- 100
```

For example :
r in N(mean=`r r`,sd = `r r_var`) and N0 = `r N`

```{r dyn3}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r, r_temporal_var = r_var,fig="print")


```


```{r parm4}
r <- -0.05
r_var <- 0.1
N <- 100
```

or another example :
r in N(mean=`r r`,sd = `r r_var`) and N0 = `r N`

```{r dyn4}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r, r_temporal_var = r_var,fig="print")


```
