---
title: "LPI test in simple situations"
author: "Romain Lorrilliere"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE,error = FALSE)
```

```{r source}
library(data.table)
library(ggplot2)
library(ggeffects)

source("c:/git/LPI_theoritical_test/functions/fun_dyn_pop.r")
source("c:/git/LPI_theoritical_test/functions/fun_samples.r")
source("c:/git/LPI_theoritical_test/functions/fun_lpi.r")


```

# Purpose

## Main scenarios

We start with some simple situations: 

- One species 
  - One population
    - Without temporal stochasticity
    - With temporal stochasticity
  - Two populations 
    - With the same value of abundance 
    - With the independent temporal variance 
- Two species
  - One population
    - Without temporal stochasticity
    - With temporal stochasticity
  - Two populations 
    - With the same value of abundance 
    - With the independent temporal stochasticity


## Generic parameters 

- dynamic model: Beverton
- number of years: 30 years
- demographic stochasticity: none, and the raw abundance will be rounded
- carrying capacity (K): 50, 100 or 500   
- growth rate (r0): 0.9, 1, 1.1
- N0: will be a proportion Pk of K in 0.1, 0.5, 0.9
Hence we model various colonization histories. 

**Parameters**

```{r param1, echo=TRUE}
v_K <- c(50,100,500)
v_r0 <- c(-0.1,0,0.1)
v_N0K <- c(0.10,0.50,0.90)

tab_param <- expand.grid(K = v_K,r=v_r0,N0K = v_N0K)
setDT(tab_param)
tab_param[,`:=`(sp = 1:.N, pop = paste0(1:.N,"_",1),N0 = K * N0K)]
tab_param <- tab_param[,.(sp,pop,K,r,N0,N0K)]


```

All combinations between the values of K, r, and N0 give `r nrow(tab_param)` scenarios. 

```{r}
print(tab_param)
```



## Sampling effect

There is no sampling; the data available in this first simple analysis are the whole temporal series. 

## Independente metrics of population dynamics

We will compare the species index of LPI and the LPI with some other metrics

**growth rate**: 

* tx = (Nt -N0) / N0
* dt = Nt/N0
* glm(N~as.factor(year)  + pop , poisson)
  glm will not use in the case of only one population par species



# One species

## One population 
 
### Without temporal stochasticity


```{r dyn1, echo=TRUE}

K <- as.list(tab_param$K)
r <- as.list(tab_param$r)
N_mean <- as.list(tab_param$N0)
N_sd <- 0
nb_sp <- nrow(tab_param)
npop <- 1
demographic_stocha <- "none"
nbyear <- 30
r_var <- 0
r_tempvar <- 0
model_dyn  <- "logistic"

d_pop  <- f_make_dyn_species(nb_sp = nb_sp,nb_pop = npop, N0_mean = N_mean,N0_var = N_sd,K=K,K_method="exact",r_mean = r, r_init_var = r_var,r_temporal_var_mean = r_tempvar,nb_year = nbyear,model_name = model_dyn,demographic_stocha = demographic_stocha, fig="print")

d_pop <- merge(d_pop,tab_param,by = c("sp","pop"))

d_pop[,Nobs := N]

```



**Trend assessment**



```{r trend1, echo=TRUE}


d_pop[,`:=`(tx = (N-N0)/N0, dt = N/N0)]



```

**LPI assessment**



```{r one_percent, echo=TRUE}

d_lpi <- assess_lpi(d_pop)

``` 


**Comparaison** 

```{r compare1}
d_lpi2 <- d_lpi[variable == "i",.(pop,year,variable,value)]
d_lpi2[,pop := gsub("all",1,pop)]
setnames(d_lpi2,"value","i")
d_trend <- d_pop[,.(pop,year,N,tx,dt,K,r,N0)]
d_compare <- merge(d_trend,d_lpi2,by=c("pop","year"))

```

**1- dt**

```{r gg2}
gg <- ggplot(d_compare,aes(x=dt,y=i,colour=year)) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```


```{r}
gg <- ggplot(subset(d_compare, pop == "18_1"),aes(x=dt,y=i,colour=year)) 
gg <- gg + geom_path()
gg <- gg + geom_point()
print(gg)

```

**2- tx**


```{r gg1}
gg <- ggplot(d_compare,aes(x=tx,y=i,colour=year)) + facet_wrap(K~r,scales="free")
gg <- gg + geom_point()
print(gg)

```
**General relation**

```{r}
md <- lm(i~ dt + I(dt^2),data = d_compare)
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",coef2,"dt² + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare,aes(x=dt,y=i))
gg <- gg + geom_smooth(method = "lm", formula = y ~ x + I(x^2),linewidth = 1.5, colour = "red", alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Overall relation between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
print(gg)


```

**Relation in declining cases*

```{r}
d_compare1 <- d_compare[dt<=1,]
md <- lm(i~ dt + I(dt^2),data = d_compare1)
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",coef2,"dt² + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare1,aes(x=dt,y=i))
gg <- gg + geom_smooth(data=d_compare,method = "lm", formula = y ~ x + I(x^2),linewidth = 1.1,linetype = "dashed", colour = "#fc9272", alpha = 0.4)
gg <- gg + geom_smooth(method = "lm", formula = y ~ x + I(x^2),linewidth = 1.5, colour = "red", alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Relation in declining cases between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
gg <- gg + coord_cartesian(xlim = c(0,1),ylim = c(0,1))
print(gg)


```




**Conclusion**

_______

### With temporal stochasticity

```{r dyn2}

r_tempvar <- 0
demographic_stocha <- "poisson"

d_pop  <- f_make_dyn_species(nb_sp = nb_sp,nb_pop = npop, N0_mean = N_mean,N0_var = N_sd,K=K,K_method="exact",r_mean = r, r_init_var = r_var,r_temporal_var_mean = r_tempvar,nb_year = nbyear,model_name = model_dyn,demographic_stocha = demographic_stocha, fig="print")

d_pop <- merge(d_pop,tab_param,by = c("sp","pop"))
d_pop[,Nobs := N]

```

Now we add a bit of temporal stochasticity with a temporal variance of growth rate parameter (r). 
We use a variance equal to `r r_tempvar`.







```{r trend2}
d_pop[,`:=`(tx = (N-N0)/N0, dt = N/N0)]

```





```{r one_percent2, echo=TRUE}
d_lpi <- assess_lpi(d_pop)

``` 
**Comparaison** 

```{r compare2}
d_lpi2 <- d_lpi[variable == "i",.(pop,year,variable,value)]
d_lpi2[,pop := gsub("all",1,pop)]
setnames(d_lpi2,"value","i")
d_trend <- d_pop[,.(pop,year,N,tx,dt,K,r,N0)]
d_compare <- merge(d_trend,d_lpi2,by=c("pop","year"))

```


**dt**

```{r gg22}
gg <- ggplot(d_compare,aes(x=dt,y=i,colour=year)) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```
**General relation**

```{r}
md <- lm(i~ dt + I(dt^2),data = d_compare)
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",coef2,"dt² + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare,aes(x=dt,y=i))
gg <- gg + geom_smooth(method = "lm", formula = y ~ x + I(x^2),linewidth = 1.5, colour = "red", alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Overall relation between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
print(gg)


```

**Relation in declining cases*

```{r}
d_compare1 <- d_compare[dt<=1,]
md <- lm(i~ dt + I(dt^2),data = d_compare1)
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",coef2,"dt² + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare1,aes(x=dt,y=i))
gg <- gg + geom_smooth(data=d_compare,method = "lm", formula = y ~ x + I(x^2),size = 1.1,linetype = "dashed", colour = "#fc9272", alpha = 0.4)
gg <- gg + geom_smooth(method = "lm", formula = y ~ x + I(x^2),linewidth = 1.5, colour = "red", alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Relation in declining cases between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
gg <- gg + coord_cartesian(xlim = c(0,1),ylim = c(0,1))
print(gg)


```


** Conclusion **


## Two populations 

### With the same value of abundance 


```{r dyn_1_2}

r_tempvar <- 0
demographic_stocha <- "none"
d_pop1  <- f_make_dyn_species(nb_sp = nb_sp,nb_pop = npop, N0_mean = N_mean,N0_var = N_sd,K=K,K_method="exact",r_mean = r, r_init_var = r_var,r_temporal_var_mean = r_tempvar,nb_year = nbyear,model_name = model_dyn,demographic_stocha = demographic_stocha, fig="print")




d_pop2 <- copy(d_pop1)
d_pop1[,g := "A"]
d_pop2[,g := "B"]


d_pop <- rbind(d_pop1,d_pop2)
d_pop <- merge(d_pop,tab_param,by = c("sp","pop"))
d_pop[,`:=`(pop = paste0(pop,"_",g),sp = paste0(sp,"_",g))]                      
d_pop[,Nobs := N]
```


```{r lpi_1_2, echo=FALSE}
d_lpi <- assess_lpi(d_pop,fig="")

``` 

We have three ways to assess the references for several populations per species: a mean way, an additional way, and a regression way. 

```{r agg_sp_2_1, echo = FALSE}
d_sp <- d_pop[,.(N_mean = mean(N)),by=.(sp,year)]

d_sp0 <- d_sp[year == 1, .(N0 = mean(N_mean)),by = sp]
d_sp <- merge(d_sp,d_sp0,by = "sp")
```

```{r}

d_sp[,`:=`(tx_mean = (N_mean-N0)/N0, dt_mean = N_mean/N0)]


```

 



```{r}

d_lpi2<- d_lpi[variable == "i",.(sp,year,variable,value)]
setnames(d_lpi2,"value","i")
d_lpi2[,sp := as.character(sp)]

d_sp[,sp := as.character(sp)]
# d_sp <- d_sp[,.(sp,year,N_mean,tx_mean,dt_mean,N_md,tx_md,dt_md)]
d_sp <- d_sp[,.(sp,year,N_mean,tx_mean,dt_mean)]
d_sp[,sp_sim := gsub("_[AB]","",sp,perl = TRUE)]

tab_param_sp <- unique(tab_param[,.(sp,K,r,N0,N0K)])
tab_param_sp[,sp := as.character(sp)]

d_trend <- merge(d_sp,tab_param_sp,by.x="sp_sim",by.y="sp")

d_compare <- merge(d_trend,d_lpi2,by=c("sp","year"))

```




```{r }
gg <- ggplot(d_compare,aes(x=dt_mean,y=i,colour=year)) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```


**General relation**

```{r}
md <- lm(i~ dt_mean + I(dt_mean^2),data = d_compare)
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",coef2,"dt² + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare,aes(x=dt_mean,y=i))
gg <- gg + geom_smooth(method = "lm", formula = y ~ x + I(x^2),linewidth = 1.5, colour = "red", alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Overall relation between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
print(gg)


```

**Relation in declining cases*

```{r}
d_compare1 <- d_compare[dt_mean<=1,]
md <- lm(i~ dt_mean + I(dt_mean^2),data = d_compare1)
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",coef2,"dt² + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare1,aes(x=dt_mean,y=i))
gg <- gg + geom_smooth(data=d_compare,method = "lm", formula = y ~ x + I(x^2),linewidth = 1.1,linetype = "dashed", colour = "#fc9272", alpha = 0.4)
gg <- gg + geom_smooth(method = "lm", formula = y ~ x + I(x^2),linewidth = 1.5, colour = "red", alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Relation in declining cases between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
gg <- gg + coord_cartesian(xlim = c(0,1),ylim = c(0,1))
print(gg)


```


** Conclusion **



### With demographic stocahiticity



```{r dyn_1_2_2}

r_tempvar <- 0
demographic_stocha <- "poisson"
d_pop1  <- f_make_dyn_species(nb_sp = nb_sp,nb_pop = npop, N0_mean = N_mean,N0_var = N_sd,K=K,K_method="exact",r_mean = r, r_init_var = r_var,r_temporal_var_mean = r_tempvar,nb_year = nbyear,model_name = model_dyn,demographic_stocha = demographic_stocha, fig="print")
d_pop1[,g := "A"]

d_pop2  <- f_make_dyn_species(nb_sp = nb_sp,nb_pop = npop, N0_mean = N_mean,N0_var = N_sd,K=K,K_method="exact",r_mean = r, r_init_var = r_var,r_temporal_var_mean = r_tempvar,nb_year = nbyear,model_name = model_dyn,demographic_stocha = demographic_stocha, fig="")
d_pop2[,g := "B"]



d_pop <- rbind(d_pop1,d_pop2)
d_pop[,sp := as.numeric(as.character((sp)))]
d_pop <- merge(d_pop,tab_param,by = c("sp","pop"))
d_pop[,`:=`(pop = paste0(pop,"_",g),sp = paste0(sp,"_",g))]                      
d_pop[,Nobs := N]
```


```{r lpi_1_2_2, echo=FALSE}
d_lpi <- assess_lpi(d_pop,fig="")

``` 

We have three ways to assess the references for several populations per species: a mean way, an additional way, and a regression way. 

```{r agg_sp_2_2, echo = FALSE}
d_sp <- d_pop[,.(N_mean = mean(N)),by=.(sp,year)]

d_sp0 <- d_sp[year == 1, .(N0 = mean(N_mean)),by = sp]
d_sp <- merge(d_sp,d_sp0,by = "sp")
```


```{r}

d_sp[,`:=`(tx_mean = (N_mean-N0)/N0, dt_mean = N_mean/N0)]


```

 

```{r}

d_lpi2<- d_lpi[variable == "i",.(sp,year,variable,value)]
setnames(d_lpi2,"value","i")
d_lpi2[,sp := as.character(sp)]

d_sp[,sp := as.character(sp)]
# d_sp <- d_sp[,.(sp,year,N_mean,tx_mean,dt_mean,N_md,tx_md,dt_md)]
d_sp <- d_sp[,.(sp,year,N_mean,tx_mean,dt_mean)]
d_sp[,sp_sim := gsub("_[AB]","",sp,perl = TRUE)]

tab_param_sp <- unique(tab_param[,.(sp,K,r,N0,N0K)])
tab_param_sp[,sp := as.character(sp)]

d_trend <- merge(d_sp,tab_param_sp,by.x="sp_sim",by.y="sp")

d_compare <- merge(d_trend,d_lpi2,by=c("sp","year"))

```

**dt**


```{r }
gg <- ggplot(d_compare,aes(x=dt_mean,y=i,colour=year)) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```


**General relation**

```{r}
md <- lm(i~ dt_mean + I(dt_mean^2),data = d_compare)
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",coef2,"dt² + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare,aes(x=dt_mean,y=i))
gg <- gg + geom_smooth(method = "lm", formula = y ~ x + I(x^2),linewidth = 1.5, colour = "red", alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Overall relation between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
print(gg)


```

**Relation in declining cases*

```{r}
d_compare1 <- d_compare[dt_mean<=1,]
md <- lm(i~ dt_mean + I(dt_mean^2),data = d_compare1)
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",coef2,"dt² + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare1,aes(x=dt_mean,y=i))
gg <- gg + geom_smooth(data=d_compare,method = "lm", formula = y ~ x + I(x^2),linewidth = 1.1,linetype = "dashed", colour = "#fc9272", alpha = 0.4)
gg <- gg + geom_smooth(method = "lm", formula = y ~ x + I(x^2),linewidth = 1.5, colour = "red", alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Relation in declining cases between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
gg <- gg + coord_cartesian(xlim = c(0,1),ylim = c(0,1))
print(gg)


```



## Ten populations 

### With demographic stochasticity


```{r setup2,eval=FALSE, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, echo = FALSE,message = FALSE,warning = FALSE,error = FALSE)
```


```{r}

r_tempvar <- 0
demographic_stocha <- "poisson"

d_pop  <- f_make_dyn_species(nb_sp = nb_sp,nb_pop = 10, N0_mean = N_mean,N0_var = N_sd,K=K,K_method="exact",r_mean = r, r_init_var = r_var,r_temporal_var_mean = r_tempvar,nb_year = nbyear,model_name = model_dyn,demographic_stocha = demographic_stocha, fig="print")

d_pop[,sp := as.character(sp)]
tab_param[,sp := as.character(sp)]
d_pop <- merge(d_pop,tab_param[,.(sp,K,r,N0,N0K)],by = c("sp"))
                   
d_pop[,Nobs := N]
```


Ajout de l'information de l'extinction local

```{r}
d_extinction_pop <- d_pop[year == 30 ,.(extinct =  N == 0),by = .(pop)]

d_extinction <- d_pop[year == 30 ,.(extinct =  as.numeric(N == 0)),by = .(pop,sp)][,.(extinct =sum(extinct)/10),by = sp]
d_extinction_year <- d_pop[,.(extinct_year =  as.numeric(N == 0)),by = .(pop,sp,year)][,.(extinct_year =sum(extinct_year)/10),by = .(sp,year)]
d_extinction[,sp := as.character(sp)]
d_extinction_year[,sp := as.character(sp)]

d_pop <- merge(d_pop,d_extinction_pop,by="pop")




```


```{r , echo=FALSE}
d_lpi <- assess_lpi(d_pop,fig="")

``` 

We have three ways to assess the references for several populations per species: a mean way, an additional way, and a regression way. 

```{r , echo = FALSE}
d_sp <- d_pop[,.(N_mean = mean(N)),by=.(sp,year)]

d_sp0 <- d_sp[year == 1, .(N0mean = mean(N_mean)),by = sp]
d_sp <- merge(d_sp,d_sp0,by = "sp")


```

```{r, echo= FALSE, eval=TRUE}

vsp <- unique(d_sp[,sp])
d_pop[,year_txt := sprintf("%02d",year)]
library(glmmTMB)
pasdetemps <- nbyear-1
for(asp in vsp){
  
    dasp <- subset(d_pop,sp == asp)
  	formule <- as.formula("N~as.factor(pop)+ year_txt")
    glm1 <- glm(formule,data=dasp,family=quasipoisson)
		sglm1 <- summary(glm1)
    coefan <- tail(matrix(coefficients(glm1)),pasdetemps)
		## coefannee vecteur des variation d'abondance par annee back transformee
    coefannee <- rbind(1,exp(coefan))
    
  
  
  md <- glmmTMB(N~year_txt + (1|pop),data= dasp, family = nbinom2)
   tab_var_abond=as.data.frame(coef(summary(md))$cond)
   setDT(tab_var_abond)
   term <- row.names(tab_var_abond)
    tab_var_abond[,`:=`(N_md = exp(Estimate),year =gsub("year_txt","",term))]
    tab_var_abond[1,`:=`(year = min(d_pop[,year]))]
    tab_var_abond[1,N_md := 1]
    tab_var_abond[,year :=as.numeric(as.character(year))]
    tab_var_abond <- tab_var_abond[,.(year,N_md)]
 tab_var_abond[,N_glm := coefannee]

  tab_var_abond[,sp := asp]

  if(asp == vsp[1]) tab_sp <- tab_var_abond else tab_sp <- rbind(tab_sp,tab_var_abond)  
}

d_sp <- merge(d_sp,tab_sp, by = c("sp","year"))
```

```{r} 
tab_param[,sp := as.character(sp)]
d_sp[,sp := as.character(sp)]
d_sp <- merge(d_sp,tab_param[,.(sp,K,r,N0,N0K)],by = c("sp"))

```




```{r}

d_sp[,`:=`(tx_mean = (N_mean-N0)/N0, dt_mean = N_mean/N0,tx_md = N_md-1, dt_md = N_md,tx_glm = N_glm-1, dt_glm = N_glm)]

#d_sp[,`:=`(tx_mean = (N_mean-N0)/N0, dt_mean = N_mean/N0)]


```



```{r}

d_lpi2<- d_lpi[variable == "i",.(sp,year,variable,value)]
setnames(d_lpi2,"value","i")
d_lpi2[,sp := as.character(sp)]

d_sp[,sp := as.character(sp)]
 d_sp <- d_sp[,.(sp,year,N_mean,tx_mean,dt_mean,N_md,tx_md,dt_md,dt_glm)]
#d_sp <- d_sp[,.(sp,year,N_mean,tx_mean,dt_mean)]
d_sp[,sp_sim := gsub("_[AB]","",sp,perl = TRUE)]

tab_param_sp <- unique(tab_param[,.(sp,K,r,N0,N0K)])
tab_param_sp[,sp := as.character(sp)]

d_trend <- merge(d_sp,tab_param_sp,by.x="sp_sim",by.y="sp")

d_compare <- merge(d_trend,d_lpi2,by=c("sp","year"))


```

Ajout d'information sur le nombre d'extinction 

```{r}


d_compare <- merge(d_compare,d_extinction,by=c("sp"))
d_compare <- merge(d_compare,d_extinction_year,by=c("sp","year"))

```


**Correlation entre toutes les méthodes de calcul**

- dt_md : glmmTMB
- dt_glm : glm
- dt_mean : variation sur la moyenne des abondances

```{r}
require(GGally)
d_compare[,extinct_cat := as.factor(cut(extinct_year,breaks=c(0,0.01,0.5,0.8,1),include.lowest= TRUE))]
ggpairs(d_compare[,.(i,dt_mean,dt_md,dt_glm,extinct_cat)],columns = 1:4, aes(color = extinct_cat),alpha = 0.5)

```


**dt**

```{r}
gg <- ggplot(d_compare,aes(x=dt_mean,y=i,colour=extinct_cat)) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```



```{r, eval = TRUE}
gg <- ggplot(d_compare,aes(x=dt_md,y=i,colour=extinct_cat)) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```

```{r, eval = TRUE}
gg <- ggplot(d_compare,aes(x=dt_glm,y=i,colour=extinct_cat)) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```


**General relation**

```{r}
md <- lm(i~ dt_mean ,data = subset(d_compare,extinct == 0))
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare,aes(x=dt_mean,y=i,group = extinct_cat, colour = extinct_cat, fill = extinct_cat))
gg <- gg + geom_smooth(method = "lm", formula = y ~ x ,linewidth = 1.5, alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Overall relation between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
print(gg)


```

**Relation in declining cases*

```{r}
d_compare1 <- d_compare[dt_mean<=1,]
md <- lm(i~ dt_mean ,data = subset(d_compare1,extinct == 0))
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```


```{r}

gg <- ggplot(d_compare1,aes(x=dt_mean,y=i,group = extinct_cat, colour = extinct_cat, fill = extinct_cat))
gg <- gg + geom_smooth(data=d_compare,method = "lm", formula = y ~ x + I(x^2),linewidth = 1.1,linetype = "dashed", colour = "#fc9272", alpha = 0.4)
gg <- gg + geom_smooth(method = "lm", formula = y ~ x ,linewidth = 1.5, alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Relation in declining cases between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
gg <- gg + coord_cartesian(xlim = c(0,1),ylim = c(0,1))
print(gg)


```



########

**without the local extinction trajectories**

```{r}
d_pop0 <- d_pop[extinct == FALSE,]

```


```{r , echo=FALSE}
d_lpi0 <- assess_lpi(d_pop0,fig="")

``` 

We have three ways to assess the references for several populations per species: a mean way, an additional way, and a regression way. 

```{r , echo = FALSE}
d_sp0 <- d_pop0[,.(N_mean = mean(N)),by=.(sp,year)]

d_sp00 <- d_sp[year == 1, .(N0mean = mean(N_mean)),by = sp]
d_sp0 <- merge(d_sp0,d_sp00,by = "sp")


```

```{r, echo= FALSE, eval=TRUE}

vsp <- unique(d_sp0[,sp])
d_pop0[,year_txt := sprintf("%02d",year)]
library(glmmTMB)
pasdetemps <- nbyear-1
for(asp in vsp){
  
    dasp <- subset(d_pop0,sp == asp)
      if(length(unique(dasp[,pop])) > 2){
  	formule <- as.formula("N~as.factor(pop)+ year_txt")
    glm1 <- glm(formule,data=dasp,family=quasipoisson)
		sglm1 <- summary(glm1)
    coefan <- tail(matrix(coefficients(glm1)),pasdetemps)
		## coefannee vecteur des variation d'abondance par annee back transformee
    coefannee <- rbind(1,exp(coefan))
    
  

  md <- glmmTMB(N~year_txt + (1|pop),data= dasp, family = nbinom2)
   tab_var_abond=as.data.frame(coef(summary(md))$cond)
   setDT(tab_var_abond)
   term <- row.names(tab_var_abond)
    tab_var_abond[,`:=`(N_md = exp(Estimate),year =gsub("year_txt","",term))]
    tab_var_abond[1,`:=`(year = min(d_pop[,year]))]
    tab_var_abond[1,N_md := 1]
    tab_var_abond[,year :=as.numeric(as.character(year))]
    tab_var_abond <- tab_var_abond[,.(year,N_md)]
 tab_var_abond[,N_glm := coefannee]
 
  tab_var_abond[,sp := asp]

  if(asp == vsp[1]) tab_sp <- tab_var_abond else tab_sp <- rbind(tab_sp,tab_var_abond)  
  } 
}

d_sp0 <- merge(d_sp0,tab_sp, by = c("sp","year"),all.x = TRUE)
```

```{r} 
tab_param[,sp := as.character(sp)]
d_sp0[,sp := as.character(sp)]
d_sp0 <- merge(d_sp0,tab_param[,.(sp,K,r,N0,N0K)],by = c("sp"))
d_nbPop <- d_pop0[,.(nb_pop = length(unique(pop))),by = "sp"]
d_sp0 <- merge(d_sp0,d_nbPop)
```




```{r}

d_sp0[,`:=`(tx_mean = (N_mean-N0)/N0, dt_mean = N_mean/N0,tx_md = N_md-1, dt_md = N_md,tx_glm = N_glm-1, dt_glm = N_glm)]

#d_sp[,`:=`(tx_mean = (N_mean-N0)/N0, dt_mean = N_mean/N0)]


```



```{r}

d_lpi20<- d_lpi0[variable == "i",.(sp,year,variable,value)]
setnames(d_lpi20,"value","i")
d_lpi20[,sp := as.character(sp)]

d_sp0[,sp := as.character(sp)]
 d_sp0 <- d_sp0[,.(sp,year,nb_pop,N_mean,tx_mean,dt_mean,N_md,tx_md,dt_md,dt_glm)]
#d_sp <- d_sp[,.(sp,year,N_mean,tx_mean,dt_mean)]
d_sp0[,sp_sim := gsub("_[AB]","",sp,perl = TRUE)]

tab_param_sp <- unique(tab_param[,.(sp,K,r,N0,N0K)])
tab_param_sp[,sp := as.character(sp)]

d_trend0 <- merge(d_sp0,tab_param_sp,by.x="sp_sim",by.y="sp")

d_compare0 <- merge(d_trend0,d_lpi20,by=c("sp","year"))


```




**Correlation entre toutes les méthodes de calcul**

- dt_md : glmmTMB
- dt_glm : glm
- dt_mean : variation sur la moyenne des abondances

```{r}
require(GGally)
ggpairs(d_compare0[,.(i,dt_mean,dt_md,dt_glm)],columns = 1:4, alpha = 0.5)

```


**dt**

```{r}
gg <- ggplot(d_compare0,aes(x=dt_mean,y=i,colour=as.factor(N0K))) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```



```{r, eval = TRUE}
gg <- ggplot(d_compare,aes(x=dt_md,y=i,colour=as.factor(N0K))) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```

```{r, eval = TRUE}
gg <- ggplot(d_compare,aes(x=dt_glm,y=i,colour=as.factor(N0K))) + facet_wrap(K~r,scales="free")
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + geom_point()
print(gg)

```


**General relation**

```{r}
md <- lm(i~ dt_mean ,data = d_compare0)
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```
```{r}

gg <- ggplot(d_compare,aes(x=dt_mean,y=i,group = N0K, colour = N0K, fill = N0K))
gg <- gg + geom_smooth(method = "lm", formula = y ~ x ,linewidth = 1.5, alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Overall relation between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
print(gg)


```

**Relation in declining cases*

```{r}
d_compare1 <- d_compare[dt_mean<=1,]
md <- lm(i~ dt_mean ,data = subset(d_compare1,extinct == 0))
smd <- summary(md)
print(smd)
tsmd <- as.data.frame(smd$coefficients)
coef1 <- round(tsmd[2,1],2)
coef2 <- round(tsmd[3,1],2)
intercept <- round(tsmd[1,1],2)
eq <- paste0("i = ",coef1,"dt + ",intercept)
#cat(eq,"\n")

R <- round(smd$adj.r.squared,2)

```


```{r}

gg <- ggplot(d_compare1,aes(x=dt_mean,y=i,group = extinct_cat, colour = extinct_cat, fill = extinct_cat))
gg <- gg + geom_smooth(data=d_compare,method = "lm", formula = y ~ x + I(x^2),linewidth = 1.1,linetype = "dashed", colour = "#fc9272", alpha = 0.4)
gg <- gg + geom_smooth(method = "lm", formula = y ~ x ,linewidth = 1.5, alpha = 0.8)
gg <- gg + geom_point()
gg <- gg + geom_point(aes(x=0,y=0), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_point(aes(x=1,y=1), size = 3, colour = "orange",alpha = 0.8)
gg <- gg + geom_abline(slope = 1,intercept=0,colour="orange",alpha=.7,linewidth=1.1)
gg <- gg + labs(x = "dt",y="indice sp du LPI", title = "Relation in declining cases between the i species index of LPI and dt",subtitle = paste(eq,"\n  R² =",R))
gg <- gg + coord_cartesian(xlim = c(0,1),ylim = c(0,1))
print(gg)


```


Exploration de cas : 

- K = 100
- r = 0.1
- N0K = 0.1

```{r}

d_pop8 <- d_pop0[sp == "8",]
d_compare8 <- d_compare0[sp == "8",]
gg <- ggplot(d_pop8,aes(x=year,y = N,colour = pop,group=pop)) + geom_point() + geom_line()
print(gg)

``` 
```{r}


gg <- ggplot(d_compare8,aes(x=year,y = i)) + geom_point() + geom_line()
gg <- gg + geom_point(aes(y=dt_mean),colour="blue") + geom_line(aes(y=dt_mean),colour="blue")
print(gg)

``` 
noir : lpi
bleu : dt_mean