---
title: "Generate the temporal abundance variation of sampled population"
author: "Romain Lorrilliere"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE,error = FALSE)
```

```{r source}
source("c:/git/LPI_theoritical_test/functions/fun_dyn_pop.r")
source("c:/git/LPI_theoritical_test/functions/fun_samples.r")
```

# One population sampled

## One population dynamic

We generate the abundance variation of one population following an exponential function. 
Thus the population abundance at t time N_t is only according to :
N(t+1) = exp(r) * N(t)
N0


```{r parm1}
r <- -0.01
r_var <- 0.0025
N <- 1000
demographic_stocha <- "poisson"

```

For example :
r in =`r r` and N0 = `r N` and with demographic stochasticity simulated thanks a pick up in a Poisson distribution. 

```{r dyn1}
l <- f_make_pop_dyn(nb_pop = 1, N0_mean = N,r_mean = r, r_temporal_var = r_var,demo_var= demographic_stocha, fig="print")


```
## Sampling

### Duration

```{r prepa_sample}

pops <- l$N
npop <- ncol(pops)
pops <- reshape2::melt(pops)
setDT(pops)
colnames(pops) <- c("year","pop","N")

```


```{r sampl_param1}
start_mean <- 5
start_var <- 0
duration_mean <- 10
duration_var <- 0
completude_mean <- 1
completude_var <- 0
proba_obs_mean <- 1
proba_obs_var <- 0

```

We model the sampled abundance variation according to a survey that starts in year `r start_mean` for a `r duration_mean` years duration. 
In this example, the survey is done yearly (completeness = `r completude_mean`), and the abundances recorded are the absolute abundance. 

```{r sampl_dyn1}
popsamples <- f_samples_pop(pops,start_mean, start_var,
                           duration_mean, duration_var,
                           completude_mean, completude_var ,
                           proba_obs_mean , proba_obs_var,
                           fig = "print")


```

### Completness



```{r sampl_param2}
completude_mean <- 0.3
completude_var <- 0


```

In other cases, the survey could not be done annually. The completeness of `r completude_mean` suggests that the survey is carried out every third year. 



```{r sampl_dyn2}
popsamples <- f_samples_pop(pops,start_mean, start_var,
                           duration_mean, duration_var,
                           completude_mean, completude_var ,
                           proba_obs_mean , proba_obs_var,
                           fig = "print")


```



### Observation probability


```{r sampl_param3}
completude_mean <- 1
completude_var <- 0
proba_obs_mean <- 10/N
proba_obs_var <- 0

```

Finally, most of the surveys on the field do not provide absolute abundance but, more often, relative abundance due to the observation probability lesser than 1.  
For example, here, the observation probability = `r proba_obs_mean`.

```{r sampl_dyn3}
popsamples <- f_samples_pop(pops,start_mean, start_var,
                           duration_mean, duration_var,
                           completude_mean, completude_var ,
                           proba_obs_mean , proba_obs_var,
                           fig = "print")


```

# Several populations 

## the populations

```{r parm2}
r <- -0.05
N_var <- 100
r_var0 <- 0.02
nb_pop <- 9
```

for example for `r nb_pop` populations:

N0 in N(mean=`r N`,sd = `r N_var`)
and r in N(mean=N(mean = `r r`,sd = `r r_var0`),sd = `r r_var`) 

```{r dyn2}
l <- f_make_pop_dyn(nb_pop = nb_pop, N0_mean = N,N0_var = N_var,r_mean = r, r_init_var = r_var0, r_temporal_var = r_var,demo_var= demographic_stocha,fig="print")


```


## Samples
```{r prepa_sample2}

pops <- l$N
npop <- ncol(pops)
pops <- reshape2::melt(pops)
setDT(pops)
colnames(pops) <- c("year","pop","N")

```


### Same samples


```{r sampl_param4}
start_mean <- 5
start_var <- 0
duration_mean <- 10
duration_var <- 0
completude_mean <- 1
completude_var <- 0
proba_obs_mean <- 1
proba_obs_var <- 0

```

We could simulate the data provide by a same protocole.  

```{r sampl_dyn4}
popsamples <- f_samples_pop(pops,start_mean, start_var,
                           duration_mean, duration_var,
                           completude_mean, completude_var ,
                           proba_obs_mean , proba_obs_var,
                           fig = "print")


```


### Different samples



```{r sampl_param5}
start_mean <- 5
start_var <- 5
duration_mean <- 10
duration_var <- 10
completude_mean <- 0.5
completude_var <- 0.5
proba_obs_mean <- 0.05
proba_obs_var <- 0.05

```

Finally, all populations are surveyed by an indepent and unique protocole. for all survey the paramters randimezed according to : 

- year of start = N(mean=`r start_mean`, sd= `r start_var`)
- duration of survey = N(mean=`r duration_mean`, sd= `r duration_var`)
- completness = N(mean=`r completude_mean`, sd= `r completude_var`) only value between 0 and 1.
- probability of observation = N(mean=`r proba_obs_mean`, sd= `r proba_obs_var`)only value between 0 and 1.

```{r sampl_dyn5}
popsamples <- f_samples_pop(pops,start_mean, start_var,
                           duration_mean, duration_var,
                           completude_mean, completude_var ,
                           proba_obs_mean , proba_obs_var,
                           fig = "print")


```